Data Visualisation in R

### 2023-11-01 MSc_Data_Vis_v3.Rmd
Created a backup PDF

---
title: "Data Visualisation in R"
author: Dr Susan Johnston [Susan.Johnston@ed.ac.uk](mailto:Susan.Johnston@ed.ac.uk)
date: "1st November 2023"
output: learnr::tutorial
runtime: shiny_prerendered
description: Practical session on data visualisation with ggplot2. Demonstrator - Sarah Dobson.
---

```{r setup, include=FALSE}
library(learnr)
library(ggplot2)
library(palmerpenguins)
library(dplyr)

@@

![](images/plots_to_learn.png){width=70%}

@@

![](images/anscombe.png){width=50%}

@@

```{r Ex1, exercise = TRUE}
summary(lm(anscombe$y1 ~ anscombe$x1))
```

@@

@@

```{r Ex3test, exercise = T}
library(palmerpenguins)
# histogram
hist(penguins$flipper_length_mm)
# plot
plot(flipper_length_mm ~ bill_length_mm, data = penguins)
# boxplot
boxplot(flipper_length_mm ~ species, data = penguins)
```

@@

![](images/Penguins.png)

@@

```{r Ex4, exercise = T}
ggplot(penguins)
```

@@

```{r Ex5, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm))
```

@@

```{r Ex7, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram() +
  labs(x = "Flipper Length (mm)", y = "Frequency", title = "Histogram of flipper length")
```

@@

```{r Ex8, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(colour = "blue", fill = "lightblue")
```

@@

```{r Ex9, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram()
```

@@

```{r Ex10, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(colour = "black")
```

@@

```{r Ex11, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_density()

@@

```{r Ex12, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram() +
  facet_wrap(~species)

@@

<details><summary>**Hint: legend label:** [click]</summary>

Specify `fill =` in `labs()`

<details><summary>**Hint: faceting:** [click]</summary>

Use `facet_grid()` with species and sex.

@@

<details><summary>**Hint: remove NAs:** [click]</summary>

Subset the data to remove the NA values e.g. `penguins_subset <- subset(penguins, !is.na(sex))`

@@

```
# 1

```
</details>

@@

```{r Ex13, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()

@@

```{r Ex14, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  stat_smooth()

@@

```{r Ex15, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  stat_smooth(method = "lm")

@@

```{r Ex16, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  stat_smooth(method = "lm")

@@

```{r Ex17, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  scale_colour_manual(values = c("blue", "red", "purple"))

@@

```{r Ex18, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point(alpha = 0.8)

@@

```{r Ex19, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point(alpha = 0.8, size = 3)

@@

```{r Ex20, exercise = T}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  scale_x_continuous(breaks = seq(35, 60, 5))

@@

```{r Ex21, exercise = T}
ggplot(penguins) +
  geom_point(aes(x = bill_length_mm, y = flipper_length_mm, colour = species))

@@

```
# 4

```
</details>

@@

```{r Ex22, exercise = T}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_density(alpha = 0.5)

@@

![](images/EDA-boxplot.png){width=70%}

@@

```{r Ex23, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin()

@@

```{r Ex24, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin() +
  geom_boxplot(notch = T)

@@

```{r Ex25, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin() +
  geom_boxplot(width = 0.4, notch = T)

@@

```{r Ex26, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin() +
  geom_boxplot(width = 0.4)

@@

```{r Ex27, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin(aes(fill = species)) +
  geom_boxplot(width = 0.4, fill = "grey90", notch = T)

@@

```{r Ex28, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_jitter()

@@

```{r Ex29, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm, colour = species)) +
  geom_boxplot() + 
  geom_jitter(width = 0.3)

@@

```{r Ex30, exercise = T}
ggplot(penguins, aes(x = species, y = flipper_length_mm, colour = species)) +
  geom_boxplot() + 
  geom_jitter(width = 0.3, alpha = 0.3) +
  theme_bw()

@@

@@

>>>>

---
title: "Data Visualisation in R"
author: "Dr Susan Johnston [Susan.Johnston@ed.ac.uk](mailto:Susan.Johnston@ed.ac.uk)"
date: "1st November 2023"
output: pdf_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(palmerpenguins)
library(dplyr)

@@

![](inst/tutorials/MSc_Data_Vis/images/plots_to_learn.png){width=70%}

@@

![](inst/tutorials/MSc_Data_Vis/images/anscombe.png){width=70%}

@@

summary(lm(anscombe$y1 ~ anscombe$x1))

@@

\pagebreak

@@

library(palmerpenguins)

# histogram
hist(penguins$flipper_length_mm)

# plot
plot(flipper_length_mm ~ bill_length_mm, data = penguins)

# boxplot
boxplot(flipper_length_mm ~ species, data = penguins)

@@

![](inst/tutorials/MSc_Data_Vis/images/Penguins.png)

@@

ggplot(penguins)

@@

ggplot(penguins, aes(x = flipper_length_mm))

@@

ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram() +
  labs(x = "Flipper Length (mm)", y = "Frequency", title = "Histogram of flipper length")

@@

ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram(colour = "blue", fill = "lightblue")

@@

@@

ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(colour = "black")

@@

ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_density()

@@

ggplot(penguins, aes(x = flipper_length_mm)) +
  geom_histogram() +
  facet_wrap(~species)

@@

@@

@@

```
\pagebreak

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  stat_smooth()

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  stat_smooth(method = "lm")

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  stat_smooth(method = "lm")

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  scale_colour_manual(values = c("blue", "red", "purple"))

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point(alpha = 0.8)

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point(alpha = 0.8, size = 3)

@@

```{r eval = F}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  scale_x_continuous(breaks = seq(35, 60, 5))

@@

ggplot(penguins) +
  geom_point(aes(x = bill_length_mm, y = flipper_length_mm, colour = species))

@@

\pagebreak

@@

@@

![](inst/tutorials/MSc_Data_Vis/images/EDA-boxplot.png)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin()

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin() +
  geom_boxplot(notch = T)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin() +
  geom_boxplot(width = 0.4, notch = T)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin() +
  geom_boxplot(width = 0.4)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_violin(aes(fill = species)) +
  geom_boxplot(width = 0.4, fill = "grey90", notch = T)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_jitter()

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm, colour = species)) +
  geom_boxplot() + 
  geom_jitter(width = 0.3)

@@

ggplot(penguins, aes(x = species, y = flipper_length_mm, colour = species)) +
  geom_boxplot() + 
  geom_jitter(width = 0.3, alpha = 0.3) +
  theme_bw()

@@

\pagebreak

### All the answers!

```{r eval = F}
# 1

@@

### Challenges Answers:

```{r eval = F}
# 4
